---
title: "Data 602 Project"
author: "Ouano, Archangelo"
date: "2023-10-24"
output: pdf_document
---

```{r, include=FALSE}
library(binom)
library(car)
library(collapsibleTree)
library(dbplyr)
library(dplyr)
library(EnvStats)
library(ggformula)
library(ggplot2)
library(gmodels)
library(htmltools)
library(ISLR)
library(knitr)
library(lawstat)
library(markdown)
library(mosaic)
library(mdsr)
library(mosaicData)
library(nycflights13)
library(olsrr)
library(plyr)
library(purrr)
library(plotly)
library(resampledata)
library(rmarkdown)
library(rpart)
library(rpart.plot)
library(rvest)
library(SDaA)
library(shiny)
library(stringi)
library(tibble)
library(tidyr)
library(tidyselect)
library(tinytex)
library(yaml)
library(shiny)
library(lubridate)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading the Traffic Accident Data
```{r}
calgary_ta = read.csv("https://raw.githubusercontent.com/Archangelo08/Data-602-Project/main/Traffic_Incidents.csv")
```

A view of the data
```{r}
head(calgary_ta, 5)
```
These are the columns in the dataset:
```{r}
colnames(calgary_ta)
```

## Data Cleaning

Converting the START_DT column from string to datetime data type:
```{r}
calgary_ta$START_DT = ymd_hms(calgary_ta$START_DT) # convert START_DT to datetime data type
```

Creating a Month and Year columns:
```{r}
calgary_ta = calgary_ta %>%
  mutate(Month = month(START_DT), Year = year(START_DT))
```

Cleaning the null values within the [QUADRANT] column
```{r}
# for null values within [QUADRANT] take the quadrant value from [INCIDENT.INFO]
calgary_ta$QUADRANT = ifelse(calgary_ta$QUADRANT == "", stri_sub(calgary_ta$INCIDENT.INFO,-2), calgary_ta$QUADRANT)

# Capitalize the extracted quadrant values
calgary_ta$QUADRANT = toupper(calgary_ta$QUADRANT)
calgary_ta = filter(calgary_ta, calgary_ta$QUADRANT == "NE" | calgary_ta$QUADRANT == "NW" | calgary_ta$QUADRANT == "SE" | calgary_ta$QUADRANT == "SW")
```

Aggregating the number of Traffic Incidents by Year, Month, Quadrant (data from 2016 to 2013)
```{r}
accidents_month = aggregate(Count~Year + Month + QUADRANT, data=calgary_ta, FUN=sum)
```
Computing the Statistics for each Quadrant (NE, NW, SE, SW)
```{r}
favstats(~Count|QUADRANT, data=accidents_month)
```

```{r}
ggplot(accidents_month, aes(x=Month, y=Count)) +
  geom_point(col="blue") +
  xlab("Month") +
  ylab("Number of Road Incedents") +
  geom_smooth(method="lm", col="red")
```

Checking the Correlation Coefficient between Month and Number of Accidents:
```{r}
cor(Count~Month, data=accidents_month)
```

Checking if the number of traffic incidents by Month can be modeled by a Normal Distribution:
```{r}
ggplot(accidents_month, aes(sample=Count)) + stat_qq(col="blue") + stat_qqline(col="red")
```


```{r}
# Create a bar plot to show the number of accidents by Quadrant
calgary_ta %>%
  group_by(QUADRANT) %>%
  summarise(TotalAccidents = sum(Count)) %>%
  ggplot(aes(x = QUADRANT, y = TotalAccidents, fill = QUADRANT)) +
  geom_bar(stat = "identity") +
  xlab("Quadrant") +
  ylab("Total Number of Accidents") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1")
```


```{r}
#Compare if there's a difference between the average number of accidents in Quadrant areas:
# Use ANOVA to compare accident counts by Quadrant
anova_result <- aov(Count ~ QUADRANT, data = calgary_ta)
summary(anova_result)
```

```{r}
#Predict the number of accidents with months as the independent variable
# Linear regression model to predict accidents based on months
model <- lm(Count ~ Month, data = accidents_month)
summary(model)

# Predict future accident counts for given months
new_data <- data.frame(Month = 13:24)  # Assuming 13 to 24 for future months
predicted_counts <- predict(model, newdata = new_data)
```
